---
title: 'Regression Models Course Project - Transmission effect on car MPG performance'
author: "Henrique Souza (github.com/htssouza)"
output:
  html_document: default
  pdf_document:
    pandoc_args:
    - +RTS
    - -K64m
    - -RTS
---

# Executive Summary

# Data Analysis

```{r echo = FALSE}
library(datasets)
suppressWarnings(library(dplyr))
library(ggplot2)
suppressWarnings(library(GGally))
library(stats)
data(mtcars)
mtcars <- mutate(mtcars, am = am + 1)
mtcars <- mutate(mtcars, am = factor(am))
levels(mtcars$am) <- c("auto", "manual")
```

MPG distribution by Transmission (only):
```{r echo = FALSE}
g <- ggplot(data = mtcars, aes(am, mpg))
g <- g + geom_violin(stat = "ydensity", color = "black", size = 2)
g <- g + xlab("Type of transmission")
g <- g + ylab("MPG density")
g
```


# Models

# Residual Analysis

# Model selection

# Conclusion

# Appendix

Plotting variable pairs (linear model fitting lines):
```{r echo = FALSE}
ggpairs(mtcars, lower = list(continuous = wrap("smooth", method = "lm")))
```




Defining our random seed and loading libraries.


Creating 1000 samples of 40 exponential each.
For all the simulations lambda will be equal to 0.2.

```{r echo = TRUE}
lambda <- 0.2
sample.size <- 40
sample.count <- 1000
```

Now lets generate our simulation data.

```{r echo = TRUE}
sample.data <- matrix(rexp(sample.size * sample.count, lambda), sample.count, sample.size)
```

Summary of our data:

```{r echo = TRUE}
str(sample.data)
```

# Sample Mean versus Theoretical Mean

Calculating sample means:

```{r echo = FALSE}
sample.means <- apply(sample.data, 1, mean)
```

In the plot below you can see: the sample mean, the theorethical mean along
with a histogram of the sample means:

```{r echo = FALSE}
g <- ggplot(data = data.frame(sample.means), aes(x = sample.means))
g <- g + labs(title = "Distribution of sample means", x = "sample mean", y = "count")
g <- g + geom_histogram(bins = 50, color = "black", fill = "lightblue")
g <- g + geom_vline(aes(xintercept=mean(sample.means), color = "sample_mean"), linetype = "dashed")
g <- g + geom_vline(aes(xintercept=(1/lambda), color = "theorethical_mean"), linetype = "dashed")
g <- g + scale_color_manual(name = "lines", values = c(sample_mean = "red", theorethical_mean = "green"))
g
```

The two reference lines are very close. The exact values are:

```{r echo = TRUE}
paste0("theorethical_mean = ", 1 / lambda)
```

```{r echo = TRUE}
paste0("sample_mean = ", mean(sample.means))
```


# Sample Variance versus Theoretical Variance

Let's compute the variance of our simulated data:

```{r echo = TRUE}
sample.vars <- apply(sample.data, 1, sd)
```

In the plot below you can see: the sample variance, the theorethical variance along
with a histogram of the sample variances:

```{r echo = FALSE}
g <- ggplot(data = data.frame(sample.vars), aes(x = sample.vars))
g <- g + labs(title = "Distribution of sample variance", x = "sample variance", y = "count")
g <- g + geom_histogram(bins = 50, color = "black", fill = "lightblue")
g <- g + geom_vline(aes(xintercept=mean(sample.vars), color = "sample_var"), linetype = "dashed")
g <- g + geom_vline(aes(xintercept=(1/lambda), color = "theorethical_var"), linetype = "dashed")
g <- g + scale_color_manual(name = "lines:", values = c(sample_var = "red", theorethical_var = "green"))
g
```

# Distribution

Plotting together the density of 1000 numbers generated using the exponential ditribution
along with a histogram of our sample means (1000 groups of 40.)

```{r echo = FALSE}
sample.exps <- rexp(sample.count, lambda)
sample.dist <- c(sample.means, sample.exps)
sample.dist.label <- factor(rep(c("mean", "exp"), each = sample.count))
sample.dist.df <- data.frame(sample = sample.dist, label = sample.dist.label)

g <- ggplot(data = sample.dist.df, aes(x = sample, fill = label, color = label))
g <- g + labs(title = "Distribution comparison", x = "Sample", y = "count")
g <- g + geom_histogram(bins = 200, alpha = 0.5)
g
```

You can see that the mean curve has the curvature of a normal.