---
title: 'Regression Models Course Project - Transmission effect on car MPG performance'
author: "Henrique Souza (github.com/htssouza)"
output:
  html_document: default
  pdf_document:
    pandoc_args:
    - +RTS
    - -K64m
    - -RTS
---

# Executive Summary

Motor Trend Magazine is interested in exploring the relationship between a set of variables and miles per gallon (MPG). They are particularly interested in the following two questions:

1. Is an automatic or manual transmission better for MPG.
1. Quantify the MPG difference between automatic and manual transmissions.

In this document we compare different models and show that *for an average car, manual transmission outperforms automatic in <strong>11%</strong> MPG efficience*.

# Data Analysis

```{r echo = FALSE, message = FALSE}
library(datasets)
library(dplyr)
library(ggplot2)
library(GGally)
library(stats)
data(mtcars)
am.levels <- c("auto", "manual")
am.auto <- as.factor(c(1, 2))
am.manual <- as.factor(c(1, 2))
levels(am.auto) <- am.levels
levels(am.manual) <- am.levels
am.auto <- am.auto[1]
am.manual <- am.manual[2]
mtcars <- mtcars %>%
  mutate(am = am + 1) %>%
  mutate(am = factor(am))
levels(mtcars$am) <- am.levels
```

How our dataset looks like:

```{r echo = FALSE, message = FALSE}
str(mtcars)
```

MPG distribution by Transmission:

```{r echo = FALSE}
g <- ggplot(data = mtcars, aes(am, mpg, fill = am))
g <- g + geom_violin(stat = "ydensity", color = "black", size = 2)
g <- g + xlab("Type of transmission")
g <- g + ylab("MPG density")
g
```

# Models

Based on the understanding of the dataset and the variables pairs plot (see Appendix),
the following models were selected for comparison:

1. Using only the Transmission.
1. Transmission and Weight.
1. Transmission, Weight and HP
1. Transmission, Weight, HP and Displacement.
1. Transmission, Weight, HP, Displacement and Gears.

Other variables are being ignored:

- Cylinder because it is a product of other variables (and it is catehorical).
- Axle Ratio (not much related to MPG efficience).
- Quarter of mile performance (related to HP).
- Engine type and Carburators which can be linked to transmission type as well.

We will be using linear models as MPG is not a binary or a count variable.

Fitting our models:
```{r echo = TRUE}
mod1 <- lm(mpg ~ am, data = mtcars)
mod2 <-  lm(mpg ~ am + wt, data = mtcars)
mod3 <-  lm(mpg ~ am + wt + hp, data = mtcars)
mod4 <-  lm(mpg ~ am + wt + hp + disp, data = mtcars)
mod5 <-  lm(mpg ~ am + wt + hp + disp + gear, data = mtcars)
```

# Model selection

Using anova function to compare models:

```{r echo = TRUE}
anova(mod1, mod2, mod3, mod4, mod5)
```

The p-values for Model 4 is already too big (>= 0.05).
We will stick with Model 3.

# Residual Analsysis

Plotting the residuals for our choosen model:

```{r echo = FALSE, message = FALSE}
mpghat <- predict(mod3)
resid <- mtcars$mpg - mpghat

g <- ggplot(data.frame(x = mtcars$mpg, y = resid), aes(x = x, y = y))
g <- g + geom_hline(yintercept = 0, size = 2)
g <- g + geom_col(size = 0.5, color = "red")
g <- g + geom_point(size = 2, color = "black")
g <- g + geom_point(size = 1, color = "white")
g <- g + geom_smooth(method = "auto")
g <- g + xlab("MPG")
g <- g + ylab("Model Residual")
g
```

# Conclusion

Using our model to compare the performance of average cars with both transmission
types we have that a *Manual* one has a better MPG performance of:

```{r echo = FALSE}
wt.mean <- mean(mtcars$wt)
hp.mean <- mean(mtcars$hp)
auto.mpg <- predict(mod3, data.frame(am = am.auto, wt = wt.mean, hp = hp.mean))
manual.mpg <- predict(mod3, data.frame(am = am.manual, wt = wt.mean, hp = hp.mean))
cat(paste(round(as.numeric(((manual.mpg - auto.mpg) / (auto.mpg)) * 100)), "%"))
```

# Appendix

Plotting all variable pairs (linear model fitting lines):
```{r echo = FALSE, message = FALSE}
ggpairs(mtcars, lower = list(continuous = wrap("smooth", method = "lm")))
```

Plotting variable pairs used in our model (linear model fitting lines):
```{r echo = FALSE, message = FALSE}
mtcars.used <- select(mtcars, mpg, am, wt, hp)
ggpairs(mtcars.used, lower = list(continuous = wrap("smooth", method = "lm")))
```

